---
title: "Harmful Agal Bloom species in OBIS"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Requirements

+ [R v4.1+]()
+ [rlang](https://CRAN.R-project.org/package=rlang)
+ [robis](https://CRAN.R-project.org/package=robis)
+ [readr](https://CRAN.R-project.org/package=readr)
+ [dplyr](https://CRAN.R-project.org/package=dplyr)
+ [sf](https://CRAN.R-project.org/package=sf)

### Installation

```
remotes::install_github("BigelowLab/habobis")
```

### Data storage

This package downloads, stores and minipulates a subset of OBIS data records for user specified species.
In an effort to make this work in your location of choice we provide you with the tools to define a data path we call `root_path`, but you can always override the default.

The easiest way to set the path is to create a hidden text file in your home directory, `~/.habobis`.  The file needs only one line of information: the path to the data set. For example, the author's has this `/mnt/ecocast/coredata/obis/hab`. This works well because it persists across R sessions.

You can also set it per session using `options(habobis_root = "/put/your/path/here")`.

### Read and fetching data - a blurred distinction.

You can explicity download a species.

```{r, message = FALSE}
suppressPackageStartupMessages({
  library(habobis)
  library(dplyr)
})
x <- fetch_species("Alexandrium affine")
glimpse(x)
```

But you can also simply **read** by species.  If you haven't already downloaded that species then it will be fetched for you.

```{r}
y <- read_species("Alexandrium andersonii")
glimpse(y)
```

### Reading multiple species

You can read more than one species at a time and bind them into one table.

```{r}
x <- read_species(c("Alexandrium affine", "Alexandrium andersonii"), bind = TRUE)
count(x, scientificname)
```

### Read as spatial objects

```{r}
library(sf)
library(leaflet)
x <- read_species(c("Alexandrium affine", "Alexandrium andersonii"), bind = TRUE, form = "sf") |>
  group_by(scientificName)

pal <- colorFactor(
  palette.colors(n = nrow(group_data(x))),
  group_data(x)$scientificName,
  na.color = "#808080"
)

leaflet(data = x) |>
  addTiles() |>
  addCircleMarkers(color = ~pal(scientificName)) |>
  addLegend(pal = pal, values = ~scientificName, opacity = 1)
```
